From OpenACC to OpenMP 4: \\ Toward Automatic Translation

### 2016-06-14 paper.tex
Fix reduction code; add reduction to loop-clauses

\begin{figure}
\begin{tabular}{rcl}
\textit{acc-directive} &$\rightarrow$& \texttt{\#pragma acc data\ }\textit{data-clauses} \\
                       &     $|$     & \texttt{\#pragma acc parallel}\texttt{\ }\textit{par-clauses} \\
                       &     $|$     & \texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses} \\
\textit{type}          &$\rightarrow$& \texttt{gang vector} \\
                       &     $|$     & \texttt{gang} \\
                       &     $|$     & \texttt{vector} \\
                       &     $|$     & \texttt{seq} \\
%                       &     $|$     & $\epsilon$ \\
\textit{data-clause}   &$\rightarrow$& \texttt{copyin(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{copyout(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{copy(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{create(}\textit{vars}\texttt{)} \\
\textit{par-clause}&$\rightarrow$&\texttt{num\_gangs(}\textit{n}\texttt{)} \\
                       &     $|$     & \texttt{vector\_length(}\textit{n}\texttt{)} \\
                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{reduction(}\textit{reductions}\texttt{)} \\
\textit{loop-clause}   &$\rightarrow$& \texttt{independent} \\
                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
\end{tabular}
\caption{OpenACC subset grammar accepted by the translation tool.  Syntax is
omitted for \textit{vars}, \textit{reductions}, and natural numbers \textit{n}.
The nonterminals \textit{data-clauses} and \textit{loop-clauses} correspond to
sequences of zero or more phrases derived from \textit{data-clause} and
\textit{loop-clause}, respectively.  In addition to the directives shown,
\texttt{\#pragma acc parallel loop} is supported, and treated as syntactic
sugar for a \texttt{loop} directive nested directly under a \texttt{parallel}
directive.}
\label{fig:grammar}
\end{figure}

@@

\begin{figure*}[t]
\begin{center}
{\footnotesize
% (a) Original
\begin{minipage}{.5\textwidth}
\texttt{%
int x, sum = 100; \\
\textcolor{blue}{\bfseries \#pragma acc loop seq private(x) reduction(+:sum)} \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} x = 100; \\
\hspace*{2em} sum += x; \\
\}
}
\end{minipage}
\hspace*{.07\textwidth}
% (b) Original
\begin{minipage}{.36\textwidth}
\texttt{%
int x; \\
\textcolor{blue}{\bfseries \#pragma acc loop seq reduction(+:x)} \\
for (int i = 0; i < 5; i++) \\
\hspace*{2em} x = 1;
}
\end{minipage} \\
% Arrows
\begin{minipage}{.5\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
\end{minipage}
\hspace*{.07\textwidth}
\begin{minipage}{.36\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
\end{minipage} \\
% (a) Transformed
\begin{minipage}{.5\textwidth}
\texttt{%
int x, sum = 100; \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} \textcolor{blue}{\bfseries int x;} \\
\hspace*{2em} x = 100; \\
\hspace*{2em} sum += x; \\
\}
}
\end{minipage}
\hspace*{.07\textwidth}
% (b) Transformed
\begin{minipage}{.36\textwidth}
\texttt{%
int x; \\
\textcolor{blue}{\bfseries int sum = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (1)} \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} \textcolor{blue}{\bfseries int x = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (2)} \\
\hspace*{2em} x = 1; \\
\hspace*{2em} \textcolor{blue}{\bfseries sum += x;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (3)} \\
\} \\
\textcolor{blue}{\bfseries x = sum;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (4)}
}
\end{minipage} \\
% Labels
\begin{minipage}{.5\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} (a)}
\end{minipage}
\hspace*{.07\textwidth}
\begin{minipage}{.36\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} (b)}
\end{minipage} \\
}
\end{center}
\caption{Elimination of \emph{private} and \emph{reduction} clauses on a sequential loop.}
\label{fig:privred}
\end{figure*}

@@

To illustrate removing a \emph{reduction} clause, consider as a concrete example
the clause \texttt{reduction(+:x)}.  There are two cases.

\textit{Case~1}.  This is perhaps the most commonly occurring use of the
\emph{reduction} clause: If the only accesses to \texttt{x} in the loop
body are in statements of the form \texttt{x++} or \texttt{x\ +=\
}\emph{expression} (where \texttt{x} is not assigned in \emph{expression}),
then the reduction clause can be removed with no change in semantics.  This is
the case in Figure~\ref{fig:privred}(a), for example.

\textit{Case~2}.  When the above condition does not hold, it is still possible
to remove a \emph{reduction} clause.  However, preserving the semantics of the
\emph{reduction} clause requires a more complex source code transformation,
illustrated in Figure~\ref{fig:privred}(b).
(1)~A ``fresh'' variable (\texttt{sum}) is declared to accumulate the result.
(2)~Since \texttt{x} is the reduction variable, a local variable \texttt{x}
is declared in the loop body with the same type as the shadowed variable
\texttt{x}, and it is initialized to the initialization value for a sum
reduction (\texttt{0}).
(3)~At the end of the loop body, the local \texttt{x} is added to the result
(\texttt{sum}).
(4)~After the loop, the accumulated result is stored in the original
\texttt{x}.

>>>>

\begin{figure}
\begin{tabular}{rcl}
\textit{acc-directive} &$\rightarrow$& \texttt{\#pragma acc data\ }\textit{data-clauses} \\
                       &     $|$     & \texttt{\#pragma acc parallel}\texttt{\ }\textit{par-clauses} \\
                       &     $|$     & \texttt{\#pragma acc loop\ }\textit{type}\texttt{\ }\textit{loop-clauses} \\
\textit{type}          &$\rightarrow$& \texttt{gang vector} \\
                       &     $|$     & \texttt{gang} \\
                       &     $|$     & \texttt{vector} \\
                       &     $|$     & \texttt{seq} \\
%                       &     $|$     & $\epsilon$ \\
\textit{data-clause}   &$\rightarrow$& \texttt{copyin(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{copyout(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{copy(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{create(}\textit{vars}\texttt{)} \\
\textit{par-clause}&$\rightarrow$&\texttt{num\_gangs(}\textit{n}\texttt{)} \\
                       &     $|$     & \texttt{vector\_length(}\textit{n}\texttt{)} \\
                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{reduction(}\textit{reductions}\texttt{)} \\
\textit{loop-clause}   &$\rightarrow$& \texttt{independent} \\
                       &     $|$     & \texttt{private(}\textit{vars}\texttt{)} \\
                       &     $|$     & \texttt{reduction(}\textit{reductions}\texttt{)} \\
\end{tabular}
\caption{OpenACC subset grammar accepted by the translation tool.  Syntax is
omitted for \textit{vars}, \textit{reductions}, and natural numbers \textit{n}.
The nonterminals \textit{data-clauses} and \textit{loop-clauses} correspond to
sequences of zero or more phrases derived from \textit{data-clause} and
\textit{loop-clause}, respectively.  In addition to the directives shown,
\texttt{\#pragma acc parallel loop} is supported, and treated as syntactic
sugar for a \texttt{loop} directive nested directly under a \texttt{parallel}
directive.}
\label{fig:grammar}
\end{figure}

@@

\begin{figure*}[t]
\begin{center}
{\footnotesize
% (a) Original
\begin{minipage}{.5\textwidth}
\texttt{%
int p, x = 100; \\
\textcolor{blue}{\bfseries \#pragma acc loop seq private(p) reduction(+:x)} \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} p = 100; \\
\hspace*{2em} x += p; \\
\}
}
\end{minipage}
\hspace*{.07\textwidth}
% (b) Original
\begin{minipage}{.36\textwidth}
\texttt{%
int x = 100; \\
\textcolor{blue}{\bfseries \#pragma acc loop seq reduction(+:x)} \\
for (int i = 0; i < 5; i++) \\
\hspace*{2em} x = 1;
}
\end{minipage} \\
% Arrows
\begin{minipage}{.5\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
\end{minipage}
\hspace*{.07\textwidth}
\begin{minipage}{.36\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} $\Downarrow$}
\end{minipage} \\
% (a) Transformed
\begin{minipage}{.5\textwidth}
\texttt{%
int p, x = 100; \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} \textcolor{blue}{\bfseries int p;} \\
\hspace*{2em} p = 100; \\
\hspace*{2em} x += p; \\
\}
}
\end{minipage}
\hspace*{.07\textwidth}
% (b) Transformed
\begin{minipage}{.36\textwidth}
\texttt{%
int x = 100; \\
\textcolor{blue}{\bfseries int sum = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (1)} \\
for (int i = 0; i < 5; i++) \{ \\
\hspace*{2em} \textcolor{blue}{\bfseries int x = 0;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (2)} \\
\hspace*{2em} x = 1; \\
\hspace*{2em} \textcolor{blue}{\bfseries sum += x;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (3)} \\
\} \\
\textcolor{blue}{\bfseries x += sum;\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ // (4)}
}
\end{minipage} \\
% Labels
\begin{minipage}{.5\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} (a)}
\end{minipage}
\hspace*{.07\textwidth}
\begin{minipage}{.36\textwidth}
\centerline{\rule[-0.5em]{2em}{0em} (b)}
\end{minipage} \\
}
\end{center}
\caption{Elimination of \emph{private} and \emph{reduction} clauses on a sequential loop.}
\label{fig:privred}
\end{figure*}

@@

To illustrate removing a \emph{reduction} clause, consider as a concrete example
\texttt{reduction(+:x)}.  There are two cases.

\textit{Case~1}.  This is perhaps the most commonly occurring use of the
\emph{reduction} clause: If the only accesses to \texttt{x} in the loop
body are in statements of the form \texttt{x++} or \texttt{x\ +=\
}\emph{expression} (where \texttt{x} does not occur in \emph{expression}),
then the reduction clause can be removed with no change in semantics.  This is
the case in Figure~\ref{fig:privred}(a), for example.

\textit{Case~2}.  When the above condition does not hold, it is still possible
to remove a \emph{reduction} clause.  However, preserving the semantics of the
\emph{reduction} clause requires a more complex source code transformation,
illustrated in Figure~\ref{fig:privred}(b).
(1)~A ``fresh'' variable (\texttt{sum}) is declared to accumulate the result.
(2)~Since \texttt{x} is the reduction variable, a local variable \texttt{x}
is declared in the loop body with the same type as the shadowed variable
\texttt{x}, and it is initialized to the initializer value for a sum
reduction (\texttt{0}).
(3)~At the end of the loop body, the local \texttt{x} is added to the result
(\texttt{sum}).
(4)~After the loop, the accumulated result is added to the original
\texttt{x}.
